Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2019-03-17 mpaper/paper.tex
I localised the bug, but still don't see anything wrong

\begin{lemma} \label{lemma:bound1}
  Let $i \in \{ 0, \dots, d \}$ and $\epsilon > 0$ be arbitrary. Furthermore,
  let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i -
  \epsilon, \lambda_i + \epsilon) \subset \mathbb{R}$ be a function with a
  codomain arbitrarily close to $\lambda_i$. Then
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  has upper and lower bounds of the form $q(\mathbf{r} \mid
  \mathbf{u})d(\mathbf{u})$, where $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
\end{lemma}
\begin{proof}
  Remember that
  \begin{align*}
    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
                                                                      &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
  \end{align*}
  by Lemma \ref{lemma:derivatives}.

  Given any covariance matrix $\mathbf{K}$, we can easily establish constant
  upper and lower bounds on $\mathbf{K}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$ by the boundedness of $c$ and continuity of the covariance
  function. Similar reasoning combined with the expressions for derivatives of
  covariance matrices in Lemma \ref{lemma:derivatives} gives constant upper and
  lower bounds on the elements of
  \[
    \left. \frac{\mathbf{\partial K}}{\partial \lambda_i} \right|_{\lambda_i =
      c(\mathbf{r}, \mathbf{u})}
  \]
  as well.

  First, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$ has constant bounds. If $i = 0$, then $\mathbf{K}|_{\lambda_i
    = c(\mathbf{r}, \mathbf{u})} = c(\mathbf{r}, \mathbf{u})\mathbf{A}$, where
  $\mathbf{A}$ is a constant matrix defined by
  \[
    \mathbf{A} = \frac{1}{\lambda_0}\mathbf{K},
  \]
  and, thus, is invertible.
  Therefore,
  \[
    \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \frac{1}{c(\mathbf{r}, \mathbf{u})}\mathbf{A}^{-1}.
  \]
  Since $\lambda_0$ cannot be zero, we can make $\epsilon$ small enough so that
  either $\lambda_0 - \epsilon > 0$ or $\lambda_0 + \epsilon < 0$, i.e., the
  values of $c(\mathbf{r}, \mathbf{u})$ are either all positive or all negative.
  Then, by continuity of $x \mapsto \frac{1}{x}$, $\mathbf{K}^{-1}|_{\lambda_i =
    c(\mathbf{r}, \mathbf{u})}$ has constant bounds.

  Let
  \[
    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{\lambda_n}{2}
    (x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n
  \]
  and $\delta = c(\mathbf{r}, \mathbf{u}) - \lambda_i$ so that $c(\mathbf{r},
  \mathbf{u}) = \lambda_i + \delta$. If $i > 0$, then
  \begin{align*}
    &k(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2\\
      - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) - S \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} (\lambda_i + \delta)(x_{j,i} - x_{k,i})^2 \\
      - \mathbbm{1}[j \ne k]\sigma^2(\lambda_i + \delta) - S \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2}(\mathbf{x}_j - \mathbf{x}_k)^\intercal \bm\Lambda (\mathbf{x}_j - \mathbf{x}_k) - \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \\
      - \frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= k(\mathbf{x}_j, \mathbf{x}_k)\exp \left( \vphantom{\frac{1}{2}} \right. -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right) \\
    &= \!\begin{multlined}[t]
      k(\mathbf{x}_j, \mathbf{x}_k) + k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp
        \left( -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. \\
      -\mathbbm{1}[j \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1 \right)
    \end{multlined}
  \end{align*}
  Hence, we can express $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ as
  $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \mathbf{K} +
  \mathbf{E}$, where
  \[
    E_{j,k} = k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp \left(
        -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. -\mathbbm{1}[j
    \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1
    \right).
  \]
  Note that $E_{j,k} \to 0$ as $\delta \to 0$ and $\delta$ can be made
  arbitrarily small via $\epsilon$. Then, since $\mathbf{K}$ is invertible, we
  can make $\delta$ small enough so that $\lVert \mathbf{K}^{-1} \rVert_\infty
  \lVert \mathbf{E} \rVert_\infty < 1$, and Lemma \ref{lemma:perturbation} shows
  the existence of $\mathbf{K}^{-1}_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ and
  gives upper and lower bounds on all of its elements.

@@

  Since $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ is bounded, so is
  $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$. Assuming that
  $\bm\Gamma$ is invertible so that $q(\mathbf{r} \mid \mathbf{u})$ exists, we
  can make $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ close enough to
  $\bm\Gamma$ to ensure that
  \[
    \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \ne 0.
  \]
  Moreover, boundedness of $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$
  gives boundedness of $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$.

  Combined with the fact that $ab > 0$ (i.e., both endpoints together are either
  positive or negative), this gives boundedness of $\det(\bm\Gamma)|_{\lambda_i
    = c(\mathbf{r}, \mathbf{u})}^{-1}$.

  % TODO: actually demonstrate this for the inverse
  Recall that
  \[
    \bm\Gamma = \Krr - \mathbf{S}\Kru = \Krr - \Kru^\intercal\Kuu^{-1}\Kru.
  \]
  We have already demonstrated that for $\mathbf{K} \in \{ \Krr, \Kru, \Kuu^{-1}
  \}$, $\lim_{\epsilon \to 0}\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  = \mathbf{K}$. We can use a general fact that if $\lim_{x \to 0} f(x) = f$,
  then $f(x) = f + g(x)$ for some function $g$ such that $\lim_{x \to 0} g(x) =
  0$ to define matrices $\Err$, $\Eru$, and $\Euu$ such that
  \begin{alignat*}{3}
    \Krr|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Krr + \Err, &&\Err \to
    \mathbf{O}_{|\mathcal{S}|}, \\
    \Kru|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kru + \Eru, \quad
    \text{and} \quad &&\Eru \to \mathbf{O}_{|\mathcal{S}|, m}, \\
    \Kuu^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kuu^{-1} + \Euu,
    &&\Euu \to \mathbf{O}_{m}.
  \end{alignat*}
  as $\epsilon \to 0$. Then, $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  = \bm\Gamma + \mathbf{E}$, where
  \begin{align*}
    \mathbf{E} &= \Err - \Kru^\intercal\Kuu^{-1}\Eru - \Kru^\intercal\Euu(\Kru + \Eru) \\
               &- \Eru^\intercal(\Kuu^{-1} + \Euu)(\Kru + \Eru) \\
               &\to \mathbf{O}_{|\mathcal{S}|}
  \end{align*}
  as $\epsilon \to 0$. Thus, Lemma \ref{lemma:perturbation} shows that
  $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ exists and provides
  constant bounds on its elements.

@@

>>>>

\begin{lemma} \label{lemma:bound1}
  Let $i \in \{ 0, \dots, d \}$ and $\epsilon > 0$ be arbitrary. Furthermore,
  let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i -
  \epsilon, \lambda_i + \epsilon) \subset \mathbb{R}$ be a function with a
  codomain arbitrarily close to $\lambda_i$. Then
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  has upper and lower bounds of the form $q(\mathbf{r} \mid
  \mathbf{u})d(\mathbf{u})$, where $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
\end{lemma}
\begin{proof}
  Remember that
  \begin{align*}
    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
                                                                      &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
  \end{align*}
  by Lemma \ref{lemma:derivatives}. Let $\mathbf{K}$ be any covariance
  matrix and
  \[
    \mathbf{A} = \frac{1}{\lambda_0} \mathbf{K}.
  \]

  First, we will show that
  \[
    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to 0 \quad \text{and}
    \quad \left. \frac{\mathbf{\partial K}}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to 0
  \]
  as $\epsilon \to 0$.
  We can easily establish constant upper and lower bounds on
  $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ by the boundedness of
  $c$ and continuity of the covariance function. Similar reasoning combined with
  the expressions for derivatives of covariance matrices in Lemma
  \ref{lemma:derivatives} gives constant upper and lower bounds on the elements
  of
  \[
    \left. \frac{\mathbf{\partial K}}{\partial \lambda_i} \right|_{\lambda_i =
      c(\mathbf{r}, \mathbf{u})}
  \]
  as well.

  % Inverse of a covariance matrix
  Now, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$ exists and
  \[
    \lim_{\epsilon \to 0} \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
      \mathbf{u})} = \mathbf{K}.
  \]
  If $i = 0$, then $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
  c(\mathbf{r}, \mathbf{u})\mathbf{A}$. Therefore\footnote{Note that since
    $\lambda_0 \ne 0$, $c(\mathbf{r}, \mathbf{u}) \ne 0$ for small enough
    $\epsilon$.},
  \begin{align*}
    \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &=
                                                               \frac{1}{c(\mathbf{r}, \mathbf{u})}\mathbf{A}^{-1} \\
                                                             &\to \frac{1}{\lambda_0}\mathbf{A}^{-1} = \frac{1}{\lambda_0} \left( \frac{1}{\lambda_0}\mathbf{K} \right)^{-1} = \mathbf{K}^{-1}
  \end{align*}
  as $\epsilon \to 0$.

  We will now show the same result for $\lambda_i > 0$. Let
  \[
    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{\lambda_n}{2}
    (x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n
  \]
  and $\delta = c(\mathbf{r}, \mathbf{u}) - \lambda_i$ so that $c(\mathbf{r},
  \mathbf{u}) = \lambda_i + \delta$, and $\lim_{\epsilon \to 0} \delta = 0$.
  Then,
  \begin{align*}
    &k(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2\\
      - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) - S \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} (\lambda_i + \delta)(x_{j,i} - x_{k,i})^2 \\
      - \mathbbm{1}[j \ne k]\sigma^2(\lambda_i + \delta) - S \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= \!\begin{multlined}[t]
      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2}(\mathbf{x}_j - \mathbf{x}_k)^\intercal \bm\Lambda (\mathbf{x}_j - \mathbf{x}_k) - \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \\
      - \frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right)
    \end{multlined} \\
    &= k(\mathbf{x}_j, \mathbf{x}_k)\exp \left( \vphantom{\frac{1}{2}} \right. -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right) \\
    &= \!\begin{multlined}[t]
      k(\mathbf{x}_j, \mathbf{x}_k) + k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp
        \left( -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. \\
      -\mathbbm{1}[j \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1 \right)
    \end{multlined}
  \end{align*}
  Hence, we can express $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ as
  $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \mathbf{K} +
  \mathbf{E}$, where $\mathbf{E}$ is defined as
  \[
    E_{j,k} = k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp \left(
        -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. -\mathbbm{1}[j
    \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1
    \right).
  \]
  By this definition,
  \[
    \lim_{\epsilon \to 0} E_{j,k} = 0.
  \]
  Then, since $\mathbf{K}$ is invertible, Lemma \ref{lemma:perturbation} shows
  the existence of $\mathbf{K}^{-1}_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ and
  gives upper and lower bounds on all of its elements.

@@

  Since
  \[
    \lim_{\epsilon \to 0} \bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \bm\Gamma,
  \]
  we also have that
  \[
    \lim_{\epsilon \to 0} \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \det(\bm\Gamma).
  \]
  Assuming that
  $\bm\Gamma$ is invertible so that $q(\mathbf{r} \mid \mathbf{u})$ exists,
  \[
    \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \ne 0
  \]
  for small enough $\epsilon$, and, thus, $\det(\bm\Gamma)^{-1}|_{\lambda_i =
    c(\mathbf{r}, \mathbf{u})}$ exists and has constant bounds.

  Recall that
  \[
    \bm\Gamma = \Krr - \mathbf{S}\Kru = \Krr - \Kru^\intercal\Kuu^{-1}\Kru.
  \]
  We have already demonstrated that for $\mathbf{K} \in \{ \Krr, \Kru, \Kuu^{-1}
  \}$,
  \[
    \lim_{\epsilon \to 0}\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
    \mathbf{K}.
  \]
  We can use a general fact that if $\lim_{x \to 0} f(x) = f$,
  then $f(x) = f + g(x)$ for some function $g$ such that $\lim_{x \to 0} g(x) =
  0$ to define matrices $\Err$, $\Eru$, and $\Euu$ such that
  \begin{alignat*}{3}
    \Krr|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Krr + \Err, &&\Err \to
    \mathbf{O}_{|\mathcal{S}|}, \\
    \Kru|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kru + \Eru, \quad
    \text{and} \quad &&\Eru \to \mathbf{O}_{|\mathcal{S}|, m}, \\
    \Kuu^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kuu^{-1} + \Euu,
    &&\Euu \to \mathbf{O}_{m}.
  \end{alignat*}
  as $\epsilon \to 0$. Then, $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  = \bm\Gamma + \mathbf{E}$, where
  \begin{align*}
    \mathbf{E} &= \Err - \Kru^\intercal\Kuu^{-1}\Eru - \Kru^\intercal\Euu(\Kru + \Eru) \\
               &- \Eru^\intercal(\Kuu^{-1} + \Euu)(\Kru + \Eru) \\
               &\to \mathbf{O}_{|\mathcal{S}|}
  \end{align*}
  as $\epsilon \to 0$. Thus, Lemma \ref{lemma:perturbation} shows that
  $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ exists and provides
  constant bounds on its elements.

@@

% TODO: a section about the algorithm
